---
title: "Dataframe Comparison and Separation"
author: "Anton"
date: "`r Sys.Date()`"
output: html_document
---

# Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Greetings!

Hi Sukanya, this is one of my most wished goals since 2019 August, to have a joint project with you, particularly am really interested in learning a ton of things from you in developmental biology especially because I find your reasoning and way of expression quite easy to understand. 


Now to not sound too British, I will get to the core purpose of sending you the markdown. Hope you find this an interesting read.

Towards the end have written some interesting points about basics of R (and python, Python took most ideas from R - for datascience needs)

# Solving The Problem (STP)

## Loading the data
Let's first load the data and then do the required operation you watned to perform on the datasets you had. 
```{r Loading_the_data}
load(STP_data)
```


## The problem from my purrview

You have two datasets which correspond to sox2 and sox3 (?) - Nah, I need to ask you once again ;( (Meow)

In these two datasets you have one column common in the both the datasets - that being Gene Names. The two datasets although have many columns different and you want to ge all the information, values stored in all the columns for Genes that are common in the two datasets and the ones that are not. 

```{r Looking_At_Dataset}
STP_data
```
Let's look at the data we have - It is a subset of Egambia dataset, which had the following description:

*This data set has been constructed from the gene expression data set accessible in the Gene Expression Omnibus (GEO) at the accession number GSE28623. Ten healthy donors (NID, non-infected donors) and 10 tubercolosis patients (TB) have been randomly selected from the full data set, and top 25 genes with the highest IQR have been selected for further analysis. Genes without an Entrez gene (EG) identifier have likewise been omitted.*

There is an easy way in R to download any data set from GEO - if you are interested would be happy to look into and discuss with you.


## Creating the problem

Now we need two data frames from this data set (dataframe = dataset for all practical purposes)

WE will create two data frames with Gene Names and EG common, but NID and TB present in either one of them but not both. Also we will sample some random 3500 (for no reason) rows in both of the data frames and find which of them are common and which are not and proceed with creating them as separate data frames that could then be stored in respective csv files.


```{r Creating_The_Problem}
STP_data %>%
  select(GENE_NAME,EG,TB) %>%
  slice_sample(n = 3500)-> STP_TB_data

STP_data %>%
  select(GENE_NAME,EG,NID) %>%
  slice_sample(n = 3500)-> STP_NID_data

```

## Proceeding with the solution

Finding the gene names in the two data frames:

This we will perform using intersect this time, as it is much cleaner and won't require to deal with boolean vectors

```{r Proceeding_with_solution}
# Let's check the column names to know which column to take intersection of 
STP_NID_data %>%
  colnames()

# The two data frames has column 'GENE_NAME' that is common
intersect(STP_NID_data["GENE_NAME"], STP_TB_data["GENE_NAME"])


```

Now we have got the 'duplicate' gene names in the two data frames. What we now will do is make two new data frames for each of STP_TB... and STP_NID..


## Data frames can be cumbersome for certain operations
We have got the duplicate gene names in either of the data frames, but we want to make it a little lighter in memeory (RAM space) and easier for operation. So we will convert them into an array or vector of characters/strings to then do downstream operations

Starting with what we did before and converting it into a better format
```{r Df_cumbersome}

intersect(STP_NID_data["GENE_NAME"], STP_TB_data["GENE_NAME"]) %>%
  unlist(use.names = F) -> STP_duplicateGeneNames

```

Now we can proceed in two different ways from here, which i really want to discuss with you:
1) We merge the two dataframes, this will make a new dataframe with 4 columns viz. GENE_NAME, EG, NID and TB. Each of the duplicate row will have values for all the four columns corressponding to the values in the two dataframes, merged. The unique rows will have 'NA' in the columns the information is present in neither of the two data frames.

2) We select duplicates from the two data frames and the merge only the rows that are duplicate - present in either of the data frames and keep the unique rows separate.


### Method 1
To proceed with this method we need to change the data types of the 'GENE_NAME' column from factor to character
This we will perform using function 'mutate', mutate the column to change it, right? I know the name could have been better...

```{r Method_1_changingToChar}

STP_TB_data %>%
  mutate(GENE_NAME = as.character(GENE_NAME)) -> STP_TB_data

STP_NID_data %>%
  mutate(GENE_NAME = as.character(GENE_NAME)) -> STP_NID_data

```
Now to merge we will use the function 'full_join'
```{r Method1_merge}
full_join(STP_NID_data, STP_TB_data)-> STP_Mthd1_MrgdDF

STP_Mthd1_MrgdDF

```
You see that the number of rows is `r nrow(STP_Mthd1_MrgdDF)` 

